Coevolution leaves a stronger imprint on interactions than on community structure

### 2017-06-06 manuscript/new.tex
revision 2

\documentclass[12pt]{article}

\usepackage[table]{xcolor}

\usepackage[hidelinks]{hyperref}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{adjustbox}

\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

% fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{textgreek}
\usepackage[greek,english]{babel}
\usepackage{amsmath}

% Code highlight and colors
\usepackage{listings}
\lstset{
  numbers=left,
  tabsize=1,
  basicstyle=\small\ttfamily,
  breaklines=true
}

\usepackage{booktabs, tabularx, longtable}
\usepackage{csquotes}
\usepackage{authblk}

% Geometry block
\usepackage[letterpaper]{geometry}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\title{Interactions retain the co-phylogenetic matching that communities lost}

\author[1,2,3]{Timothée~Poisot}
\author[1]{Daniel B.~Stouffer}
\affil[1]{Centre for Integrative Ecology, School of Biological Sciences,
University of Canterbury, Christchurch, New Zealand}
\affil[2]{Université de Montréal, Département de Sciences Biologiques, Montréal,
Canada}
\affil[3]{Québec Centre for Biodiversity Sciences, Montréal, Canada}

@@

\maketitle

\begin{abstract}
  Both species and their interactions are affected by changes that occur
  at evolutionary time-scales, and these changes shape both ecological
  communities and their phylogenetic structure. That said, extant
  ecological community structure is contingent upon random chance,
  environmental filters, and local effects. It is therefore unclear how
  much ecological signal local communities should retain. Here we show
  that, in a host--parasite system where species interactions vary
  substantially over a continental gradient, the ecological significance
  of individual interactions is maintained across different scales.
  Notably, this occurs despite the fact that observed community variation
  at the local scale frequently tends to weaken or remove community-wide
  phylogenetic signal. When considered in terms of the interplay between
  community ecology and coevolutionary theory, our results demonstrate
  that individual interactions are capable and indeed likely to show a
  consistent signature of past evolutionary history even when woven into
  communities that do not.
\end{abstract}

% pandoc-xnos: macro to create a caption without a prefix
\makeatletter
\long\def\@makenoprefixcaption#1#2{
  \vskip\abovecaptionskip
  \sbox\@tempboxa{#2}
  \ifdim \wd\@tempboxa >\hsize
    #2\par
  \else
    \global \@minipagefalse
    \hb@xt@\hsize{\hfil\box\@tempboxa\hfil}
  \fi
  \vskip\belowcaptionskip}
\makeatother

% pandoc-fignos: save original macros
\makeatletter
\let\@oldmakecaption=\@makecaption
\let\oldthefigure=\thefigure
\let\oldtheHfigure=\theHfigure
\makeatother

% pandoc-fignos: environment disables figure caption prefixes
\makeatletter
\newcounter{figno}
\newenvironment{no-prefix-figure-caption}{
  \let\@makecaption=\@makenoprefixcaption
  \renewcommand\thefigure{x.\thefigno}
  \renewcommand\theHfigure{x.\thefigno}
  \stepcounter{figno}
}{
  \let\thefigure=\oldthefigure
  \let\theHfigure=\oldtheHfigure
  \let\@makecaption=\@oldmakecaption
  \addtocounter{figure}{-1}
}
\makeatother

@@

Ecological interactions often exert important selective pressures on the
species involved. For example, the phenologies of lodgepole pines and
red crossbills respond spatially to the presence of squirrels (Benkman
et al. 2003). Likewise, palm species undergo changes in seed morphology
in response to the extinction of birds dispersing their seeds (Galetti
et al. 2013; Johnson et al. 2017). Interactions can be lost, too, when
phenologies of the species involved shift (Rafferty et al. 2015).
Interactions are, in fact, so important that the existence of a species
has been inferred by the fact that another species bore traits that
matched no other known species: Kritsky (1991) relates the discovery of
the moth \emph{Xanthopan morganii}, with a proboscis famously over a
foot long, which Darwin predicted would exist based solely on the
phenology of local plant \emph{Angraecum sesquipedale}. In addition,
interactions and the emergent structures they define are distributed in
similar ways across communities at both large or small scales (Jordano
et al. 2003). Together, these observations suggest that much ecological
structure could be the end result of (co)evolutionary dynamics between
species (Eklöf et al. 2012; Stouffer et al. 2012). Unfortunately,
although the evolutionary dynamics of pairs of interacting species have
been well described at macro-evolutionary (Van Valen 1973) and
micro-evolutionary (Gandon et al. 2008) timescales, most attempts to
understand how they cascade up to the levels of diversity of both
species and interactions found within empirical communities have been
inconclusive (Hembry et al. 2014). This suggests that these
well-describe mechanisms may not confer substantial predictive power
when examined at scales of organization larger than the pairwise
interaction.

Historically, the evidence for shared evolutionary history in
taxonomically diverse communities relied on the quantification of the
degree of matching between the phylogenies of two sets of interacting
organisms, accounting for the distributions of interactions across the
phylogeny (Legendre et al. 2002). This notion builds on the century-old
idea that extant species interact in a way similar to the way their
ancestors did (Fahrenholz 1913; Guimarães Jr et al. 2011; Nuismer et al.
2013). Note that testing these assumptions is related to, but markedly
more restrictive than, testing for phylogenetic conservatism of species'
interactions (Rezende et al. 2007; Eklöf et al. 2012). This is because
of additional, higher-order constraints related to the shape of both
trees at \emph{all} depths (Cavender-Bares et al. 2009; Mouquet et al.
2012), because ancestral evolutionary innovations have a high
phylogenetic inertia, and they carry forward to extant taxa (Desdevises
et al. 2003; Diniz-Filho \& Bini 2008; Vale \& Little 2010). In a way,
the true measure of phylogenetic signal of interactions should depend
not only on how they are conserved within the tree of the species
establishing them (\emph{e.g.} parasites or pollinators), but also how
these interactions at matched to the tree of the species receiving them
(\emph{e.g.} hosts or plants). This is true whether or not the species
complex are coevolving: in fact, neutral interactions can yield
perfectly matching co-phylogenies (Poisot 2015). Consequently, many of
the systems that have been described as exhibiting significant
phylogenetic structure of interactions ultimately deviate from this last
constraint, and this can occur for a variety of factors that stem from
how other species evolved and established, lost, or maintained
interactions throughout their joint evolutionary history. Nonetheless,
detecting matching phylogenies for interacting clades indicates that
their shared evolutionary history is long standing and is therefore
suggestive that their extant ecological structure is an outcome of
ancestral constraints and/or co-adaptation (Nuismer \& Harmon 2015).

It is important to note further that discovering matching phylogenies
does not mean that coevolutionary dynamics---\emph{sensu} Thompson
(1999)---took place at any time. In fact, coevolution is not expected to
necessarily result in matching phylogenies nor are matching phylogenies
only produced through coevolution (Poisot 2015). It follows that
community-level measures of phylogenetic signal, while they \emph{do}
quantify how closely interactions are a product of phylogeny, do not
allow us to draw conclusions on coevolution. Nevertheless,
interaction-level measures are useful, in that, when expressed as the
contribution of interactions to the overall signal, they allow us to
\emph{compare} the importance of interactions across replicated
communities. Communities from the same regional pool vary because (i)
the local species pool is at best a subset of the regional species pool
and (ii) the local interactions are at best a subset of the interactions
in the regional community (Poisot et al. 2015). This implies that (i)
the phylogenetic signal in the regional pool will be different from the
signal in the local communities, and (ii) the phylogenetic signal across
local communities will differ. Species sampling and variability of
interactions, however, does no predict (i) how the phylogenetic signal
of pairwise interactions is kept or lost at the scale of the whole
community nor (ii) whether or not this variability is related to changes
in the amount of phylogenetic signal that can be detected locally.

@@

\emph{Continental} interaction data consists of the aggregated
\enquote{metanetwork} which includes all documented interactions between
species from the regional species pool (Poisot et al. 2012). For every
location, we further define two scales of analysis.

@@

Second, \emph{local} interaction data are the actual observations at
this location: the identity of the present species, and the way they
interact. In addition to capturing the dissimilarity of species
composition across sites, this allows measuring the effect of
interaction turnover across space. The local and regional networks
always include the same species, but the local network has only a subset
(or, at most, an exact match) of the interactions in the regional
network.

@@

The spatial consistency of an interaction \(S_{ij}\) between species
\(i\) and \(j\) is therefore the ratio between the the number of
locations in which they were observed to interact (\(L_{ij}\)) and the
number of locations in which both were observed to be present
(\(C_{ij}\)). Because \(L_{ij} \in [0,C_{ij}]\), this measure takes
values in \([0,1]\). Larger values reflect high spatial consistency.
Note that although they are reported as 0 (\emph{i.e.} having no
interactions), we actually have no information about species pairs that
have never co-occurred; this is a common, but hard-to-correct-for,
feature of spatially replicated datasets in which species occurrence
varies (Morales-Castilla et al. 2015). Therefore, the only values of
\(S_{ij}\) can be properly estimated are those for species pairs that
have been observed to \emph{co-occur} at least once.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figures/figure1.pdf}
\caption{Spatial distribution of co-phylogenetic matching across the 51
sites. For each location, we indicate whether or not the structure of
regional and local interaction networks is consistent with phylogenetic
congruence. The colour of the circle corresponds to regionally
significant or non-significant (black and grey, respectively) while the
colour of the symbol within corresponds to locally significant or
non-significant (black and grey, respectively).\label{fig:maps}}
\end{figure}

@@

Splitting the datasets at the continental, regional, and local levels
delineates clear quantitative predictions. At the regional scale, one
can expect community assembly to promote the co-occurrence of
evolutionarily linked species pairs -- \emph{i.e.}, a host and a
parasite from lineages that interact will tend to co-occur more often
because the parasites are filtered to be present in sites where they can
find hosts. Under this situation, we expect that regional networks will
have a high degree of phylogenetic matching (because they account for
the information on potential species interactions); we do in addition
expect that their phylogenetic signal will be larger than what is found
in the continental network, since the latter represents a somewhat
artefactual agglomeration of species pairs that do not co-occur. The
opposite situation (a relatively lower phylogenetic matching) would
therefore be suggestive of a weaker selection for the co-occurrence of
evolutionarily tied species pairs.

At the local scale, if interactions between species at matching
phylogenetic positions are conserved, we would expect both a similar or
higher level of phylogenetic matching between the local and the regional
scale, and a positive relationship between the frequency of interaction
and its overall importance for phylogenetic matching (interactions with
a strong phylogenetic signal happen more often). On the contrary, if
local assembly proceeds largely independently from the co-evolutionary
history, the relative level of phylogenetic matching in local networks
should be the same as in the regional networks (through a sampling
effect from the distribution of interaction-level contribution to
cophylogenetic matching), but the frequency of interactions should bear
no relationship to their importance in overall matching.

@@

As host-macroparasite interactions are hypothesized to be ecologically
constrained, as a result of their being evolutionary conserved (Combes
2001), the congruence observed at the continental level sets the
baseline for what would be expected in local communities. Of course, if
ecological mechanisms (such as filtering) reduce co-phylogenetic
matching, we should detect this signal at the continental scale but not
locally. Out of 51 sites, our \emph{PACo} analysis indicates that 35
show no signal of co-phylogenetic matching at all, 11 show significant
co-phylogenetic matching when using the regional interactions, and 12
show significant co-phylogenetic matching using the local interactions
(see \emph{Supp. Mat. 1} for network-level significance values;
\autoref{maps}). These results support the idea that macro-evolutionary
processes, such as co-diversification, can have consequences at the
macro-ecological level but may not in fact be detectable at finer
spatial scales.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figures/figureLocReg.pdf}
\caption{The regional and local networks show the same relative amount
of co-phylogenetic matching. The values presented are the z-scores of
the PACo statistic for the entire network, with the 1:1 relationship
indicated by the solid line.\label{fig:relative}}
\end{figure}

When we compared the relative degree of co-phylogenetic matching in the
local and regional communities (\autoref{relative}), we see that the
relationship between the two is approximately linear (95\% confidence
interval for the correlation coefficient \(0.914\)--\(0.971\)). This
fits with the hypothesis of local networks being assembled by a random
sampling from regional networks: if interactions between species at
matching positions in both trees are maintained by the same set of
drivers, then this should be reflected in the local networks by a higher
degree of cophylogenetic matching.

@@

\begin{no-prefix-figure-caption}

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figures/figure4.pdf}
\caption{Distribution of co-phylogenetic matching at the network and
interaction levels. \textbf{a}, Networks that have lower co-phylogenetic
matching at the local or regional level are composed of interactions
that on average contribute little to co-phylogenetic matching at the
continental scale. co-phylogenetic matching is presented relatively to
the continental level co-phylogenetic matching. Dashed lines are a cubic
smoothing spline, and the two levels of the same networks are linked by
solid grey lines. \textbf{b}, Overall, interactions observed at the
local, regional, and continental scale have roughly equivalent
contributions to co-phylogenetic matching. Probability density was
smoothed using a Gaussian kernel density estimator. Raw probability
densities are shown as semi-transparent bars.}\label{fg:contributions}
\end{figure}

\end{no-prefix-figure-caption}

On the other hand, system-level differences say little about the
behavior of individual interactions. Despite the fact most
coevolutionary mechanisms act at the interaction level (Thompson 1999),
most \emph{measures} of it are expressed at the community level. We
observe here that networks with interactions that are important for
co-phylogenetic matching at the continental scale are also important for
co-phylogenetic matching at the local and regional scales as well
(\(\rho = 0.95\); \autoref{contributions}A). Intriguingly, we also find
that the distribution of individual interactions' contributions to
co-phylogenetic matching is strongly conserved, regardless of the scale
at which the interactions are quantified (\autoref{contributions}B).
Because interactions differ between each other in terms of their total
contribution to co-phylogenetic matching, this implies that their
distribution across networks (\emph{i.e.} whether the local network
contains a sample of strongly contributing, or weakly contributing,
interactions) is what actually drives differences in overall
co-phylogenetic matching. As such, network-level co-phylogenetic
matching emerges directly from the properties of interactions and is not
a property of the network itself.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{figures/figure3.pdf}
\caption{Spatial consistency of an interaction and its contribution to
co-phylogenetic matching. Note that because \emph{PACo} gives low scores
to interactions with a strong contribution to co-phylogenetic matching,
the y axis is reversed. Spatial consistency is defined as the
probability of observing an interaction between two species given that
they were observed to co-occur. Although statistically significant,
there was no biologically meaningful relationship between spatial
consistency and an interaction's importance for co-phylogenetic matching
in the continental network (\(R^2 \approx 0.01\), \(\rho = -0.1\),
\(p \leq 10^{-5}\)).\label{fig:consistency}}
\end{figure}

Beyond their contribution to co-phylogenetic matching, interactions also
ultimately differ in how frequently they vary when the species involved
co-occur (Carstensen et al. 2014; Olito \& Fox 2015; Trøjelsgaard et al.
2015). This can happen, for example, when one of the partners is able to
forage for optimal resources (Betts et al. 2015). Once more, the
literature on host-parasite interactions assumes that the reason why
some interactions are more frequent is because they reflect a
significant past history of coevolution (Guimaraes et al. 2007; 2010);
that is, the ecological constraints emerge from evolutionary
conservatism. Using a weighted Pearson's correlation between the
interaction frequency, interaction contribution to co-phylogenetic
matching, and the number of observations of each interaction as the
weight, we observe that this is marginally true (\(\rho \approx -0.11\).
\(t \approx -5.09\) with weights; \(\rho \approx -0.10\),
\(t \approx -4.6\) without; both significant at \(\alpha = 0.05\);
\autoref{consistency}). Recall that the \emph{negative} correlation here
arises from the fact that high interaction-level values in PACo means
\emph{low} contribution to co-phylogenetic signal. Nevertheless, the
significance of this result ought to be tempered by the fact that the
\(R^2\) of both regressions is close to \(0.01\). Consequently, the
association between spatial consistency and contribution to
co-phylogenetic signal, while statistically significant, explains so
little variance of either quantities that it is likely of negligible
biological importance. This implies that the spatial consistency of an
interaction does not necessarily reflect its evolutionary past, but
rather (possibly) extant ecological processes.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figures/figure2.pdf}
\caption{The contribution to co-phylogenetic matching of the interaction
between two species is maintained across scales. For every site (ranked
on the x axis), we show the Pearson's correlation between
interaction-level values of co-phylogenetic matching in the continental
network and the same in the local network. The size of each point is
proportional to the size of the network, and correlations for all sites
are significant at \(\alpha = 0.05\) except for those falling in the
grey shaded area.\label{fig:scales}}
\end{figure}

Ultimately, co-phylogenetic matching varies across scale because of the
simultaneous variation of species' interactions \emph{and} communities'
phylogenetic tree structure. In a system characterised by substantial
turnover, we would expect the contribution of each separate interaction
to differ across scales as well. Instead, we observe here that
interactions that contribute strongly to co-phylogenetic matching at the
continental scale \emph{also} show a significant tendency to contribute
strongly at the local (\(p < 0.05\) for positive correlations in 48 out
of 51 networks) and regional (in 47 out of 51 networks), and this
observation is independent of network-wide co-phylogenetic matching
(\autoref{scales}). Remarkably, this result implies that the remnants of
co-phylogenetic inertia are still locally detectable in \emph{individual
interactions} even though shared evolutionary history regularly fails to
leave its imprint on most local networks.

@@

\hypertarget{ref-coel17nbp}{}
\textbf{Coelho et al.} (2017). Neutral biogeography of phylogenetically
structured interaction networks. \emph{Ecography.}

@@

>>>>

\documentclass[11pt,serif]{article}

\input{.plmt/plmt.tex}

\geometry{margin=2.5cm}

\usepackage[right]{lineno}
\linenumbers
\usepackage{endfloat}

@@

\thispagestyle{empty}

{\bfseries\huge Interactions retain the co-phylogenetic matching that communities lost}

\vskip 5em

\begin{flushleft}
\href{http://orcid.org/0000-0002-0735-5184}{Timothée Poisot}
~\emph{1,2,3}
\quad
\mbox{Daniel B. Stouffer}
~\emph{1}
\end{flushleft}
\bigskip

\small
\textbf{1}:~Centre for Integrative Ecology, School of Biological Sciences,
University of Canterbury, Christchurch, New Zealand;\quad\textbf{2}:~Université de Montréal, Département de Sciences Biologiques, Montréal,
Canada;\quad\textbf{3}:~Québec Centre for Biodiversity Sciences, Montréal, Canada\normalsize
\vfill

Both species and their interactions are affected by changes that occur
at evolutionary time-scales, and these changes shape both ecological
communities and their phylogenetic structure. That said, extant
ecological community structure is contingent upon random chance,
environmental filters, and local effects. It is therefore unclear how
much ecological signal local communities should retain. Here we show
that, in a host--parasite system where species interactions vary
substantially over a continental gradient, the ecological significance
of individual interactions is maintained across different scales.
Notably, this occurs despite the fact that observed community variation
at the local scale frequently tends to weaken or remove community-wide
phylogenetic signal. When considered in terms of the interplay between
community ecology and coevolutionary theory, our results demonstrate
that individual interactions are capable and indeed likely to show a
consistent signature of past evolutionary history even when woven into
communities that do not.
\vskip 1em
{\small{\bfseries Keywords:}
species interactions  - host-parasites  - phylogenetic congruence  - network cophylogeny  - community phylogenetics \vskip 4em
}

\vfill

{\small%
{\ccby}\\
{\emph{This work is licensed under a %
Creative Commons Attribution 4.0 Unported License.}}%
\\
Correspondence to Timothée Poisot -- \texttt{timothee.poisot@umontreal.ca}\newline
\texttt{2017-03-02T00:00:00.000Z}
}%

\cleardoublepage

\doublespacing

@@

Ecological interactions often exert important selective pressures on the
species involved. For example, the phenologies of lodgepole pines and
red crossbills respond spatially to the presence of squirrels (Benkman
et al. 2003). Likewise, palm species undergo changes in seed morphology
in response to the extinction of birds dispersing their seeds (Galetti
et al. 2013; Johnson et al. 2017). Interactions can be lost, too, when
phenologies of the species involved shift (Rafferty et al. 2015).
Interactions are, in fact, so important that the existence of a species
has been inferred by the fact that another species bore traits that
matched no other known species: Kritsky (1991) relates the discovery of
the moth \emph{Xanthopan morganii}, with a proboscis famously over a
foot long, which Darwin predicted would exist based solely on the
phenology of local plant \emph{Angraecum sesquipedale}. In addition,
interactions and the emergent structures they define are distributed in
similar ways across communities at both large or small scales (Jordano
et al. 2003). Together, these observations suggest that much ecological
structure could be the end result of (co)evolutionary dynamics between
species (Eklöf et al. 2012; Stouffer et al. 2012). Unfortunately,
although the evolutionary dynamics of pairs of interacting species have
been well described at macro-evolutionary (Van Valen 1973) and
micro-evolutionary (Gandon et al. 2008) timescales, most attempts to
understand how they cascade up to the levels of diversity of both
species and interactions found within empirical communities have been
inconclusive (Hembry et al. 2014). This suggests that these
well-described mechanisms may not confer substantial predictive power
when examined at scales of organization larger than the pairwise
interaction.

Historically, the evidence for shared evolutionary history in
taxonomically diverse communities relied on the quantification of the
degree of matching between the phylogenies of two sets of interacting
organisms, accounting for the distributions of interactions across the
phylogeny (Legendre et al. 2002). This notion builds on the century-old
idea that extant species interact in a way similar to the way their
ancestors did (Fahrenholz 1913; Guimarães Jr et al. 2011; Nuismer et al.
2013). Note that testing these assumptions is related to, but markedly
more restrictive than, testing for phylogenetic conservatism of species'
interactions (Rezende et al. 2007; Eklöf et al. 2012). This is because
of additional, higher-order constraints related to the shape of both
trees at \emph{all} depths (Cavender-Bares et al. 2009; Mouquet et al.
2012), because ancestral evolutionary innovations have a high
phylogenetic inertia, and they carry forward to extant taxa (Desdevises
et al. 2003; Diniz-Filho \& Bini 2008; Vale \& Little 2010). In a way,
the true measure of phylogenetic signal of interactions should depend
not only on how they are conserved within the tree of the species
establishing them (\emph{e.g.} parasites or pollinators), but also how
these interactions at matched to the tree of the species receiving them
(\emph{e.g.} hosts or plants). This is true whether or not the species
complex are coevolving: in fact, neutral interactions can yield
perfectly matching co-phylogenies (Poisot 2015). Consequently, many of
the systems that have been described as exhibiting significant
phylogenetic structure of interactions ultimately deviate from this last
constraint, and this can occur for a variety of factors that stem from
how other species evolved and established, lost, or maintained
interactions throughout their joint evolutionary history (in addition to
the fact that species may have different impacts on one another,
\emph{i.e.} the effect of any interaction on the community-level
co-phylogenetic signal is not expected to be equal). Nonetheless,
detecting matching phylogenies for interacting clades indicates that
their shared evolutionary history is long standing and is therefore
suggestive that their extant ecological structure is an outcome of
ancestral constraints and/or co-adaptation (Nuismer \& Harmon 2015).

It is important to note further that discovering matching phylogenies
does not mean that coevolutionary dynamics---\emph{sensu} Thompson
(1999)---took place at any time. In fact, coevolution is not expected to
necessarily result in matching phylogenies nor are matching phylogenies
only produced through coevolution (Poisot 2015). It follows that
community-level measures of phylogenetic signal, while they \emph{do}
quantify how closely interactions are a product of phylogeny, do not
allow us to draw conclusions on coevolution. Nevertheless,
interaction-level measures are useful in that, when expressed as the
contribution of interactions to the overall signal, they allow us to
\emph{compare} the importance of interactions across replicated
communities. Communities from the same regional pool vary because (i)
the local species pool is at best a subset of the regional species pool
and (ii) the local interactions are at best a subset of the interactions
in the regional community (Poisot et al. 2015). This implies that (i)
the phylogenetic signal in the regional pool will be different from the
signal in the local communities, and (ii) the phylogenetic signal across
local communities will differ. Species sampling and variability of
interactions, however, do no predict (i) how the phylogenetic signal of
pairwise interactions is kept or lost at the scale of the whole
community nor (ii) whether or not this variability is related to changes
in the amount of phylogenetic signal that can be detected locally.

@@

\emph{Continental} interaction data consists of the aggregated
``metanetwork'' which includes all documented interactions between
species from the regional species pool (Poisot et al. 2012). For every
location, we further define two scales of analysis.

@@

Second, \emph{local} interaction data are the actual observations at
this location: the identity of the species present, and the way they
interact. In addition to capturing the dissimilarity of species
composition across sites, this allows measuring the effect of
interaction turnover across space. The local and regional networks
always include the same species, but the local network has only a subset
(or, at most, an exact match) of the interactions in the regional
network.

@@

The spatial consistency of an interaction \(S_{ij}\) between species
\(i\) and \(j\) is therefore the ratio between the the number of
locations in which they were observed to interact (\(L_{ij}\)) and the
number of locations in which both were observed to be present
(\(C_{ij}\)). Because \(L_{ij} \in [0,C_{ij}]\), this measure takes
values in \([0,1]\) where larger values reflect high spatial
consistency. Note that although they are reported as 0 (\emph{i.e.}
having no interactions), we actually have no information about species
pairs that have never co-occurred; this is a common, but
hard-to-correct-for, feature of spatially replicated datasets in which
species occurrence varies (Morales-Castilla et al. 2015). Therefore, the
only values of \(S_{ij}\) can be properly estimated are those for
species pairs that have been observed to \emph{co-occur} at least once.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figure/maps-1.pdf}
\caption{Spatial distribution of co-phylogenetic matching across the 51
sites. For each location, we indicate whether or not the structure of
regional and local interaction networks is consistent with phylogenetic
congruence. Locations where the local network shows a significant
co-phylogenetic structure are marked with triangles. Locaitions where
the regional network shows a significant co-phylogenetic structure are
in black.\label{fig:maps}}
\end{figure}

@@

Splitting the datasets at the continental, regional, and local levels
delineates clear quantitative predictions. At the regional scale, one
can expect community assembly to promote the co-occurrence of
evolutionarily linked species pairs -- \emph{i.e.}, a host and a
parasite from lineages that interact will tend to co-occur more often
because the parasites are filtered to be present in sites where they can
find hosts. Under this situation, we expect that regional networks will
have a high degree of phylogenetic matching (because they account for
the information on potential species interactions); we do in addition
expect that their phylogenetic signal will be larger than what is found
in the continental network, since the latter represents a somewhat
artifactual agglomeration of species pairs that do not co-occur. The
opposite situation (a relatively lower phylogenetic matching) would
therefore be suggestive of a weaker selection for the co-occurrence of
evolutionarily tied species pairs.

At the local scale, if interactions between species at matching
phylogenetic positions are conserved, we would expect both a similar or
higher level of phylogenetic matching between the local and the regional
scale, and a positive relationship between the frequency of interaction
and its overall importance for phylogenetic matching (interactions with
a strong phylogenetic signal happen more often). On the contrary, if
local assembly proceeds largely independently from the coevolutionary
history, the relative level of phylogenetic matching in local networks
should be the same as in the regional networks (through a sampling
effect from the distribution of interaction-level contribution to
co-phylogenetic matching), but the frequency of interactions should bear
no relationship to their importance in overall matching.

@@

As host-macroparasite interactions are hypothesized to be ecologically
constrained due to their being evolutionary conserved (Combes 2001), the
congruence observed at the continental level sets the baseline for what
would be expected in local communities. Of course, if ecological
mechanisms (such as filtering) reduce co-phylogenetic matching, we
should detect this signal at the continental scale but not locally. Out
of 51 sites, our \emph{PACo} analysis indicates that 35 show no signal
of co-phylogenetic matching at all, 11 show significant co-phylogenetic
matching when using the regional interactions, and 12 show significant
co-phylogenetic matching using the local interactions (see \emph{Supp.
Mat. 1} for network-level significance values;
\xrefname{table}\cref{tbl:contingency}; \xrefname{fig.}\cref{fig:maps}).
These results support the idea that macro-evolutionary processes, such
as co-diversification, can have consequences at the macro-ecological
level but may not in fact be detectable at finer spatial scales.

\begin{longtable}[]{@{}lrr@{}}
\caption{Contingency table showing the number of networks with
(non-)significant local and regional co-phylogenetic signal.
\label{tbl:contingency}}\tabularnewline
\toprule
& Regional non-significant & Regional significant\tabularnewline
\midrule
\endfirsthead
\toprule
& Regional non-significant & Regional significant\tabularnewline
\midrule
\endhead
Local non-significant & 35 & 4\tabularnewline
Local significant & 5 & 7\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{figure/localregional-1.pdf}
\caption{The regional and local networks show the same relative amount
of co-phylogenetic matching. The values presented are the z-scores of
the PACo statistic for the entire network, with the 1:1 relationship
indicated by the solid line.\label{fig:relative}}
\end{figure}

When we compared the relative degree of co-phylogenetic matching in the
local and regional communities (\xrefname{fig.}\cref{fig:relative}), we
see that the relationship between the two is approximately linear (95\%
confidence interval for the correlation coefficient
\(0.914\)--\(0.971\)). This fits with the hypothesis of local networks
being assembled by a random sampling from regional networks: if
interactions between species at matching positions in both trees are
maintained by the same set of drivers, then this should be reflected in
the local networks by a higher degree of co-phylogenetic matching.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figure/contributions-1.pdf}
\caption{Distribution of co-phylogenetic matching at the network and
interaction levels. \textbf{a}, Networks that have lower co-phylogenetic
matching at the local or regional level are composed of interactions
that on average contribute little to co-phylogenetic matching at the
continental scale. Co-phylogenetic matching is presented relatively to
the continental level co-phylogenetic matching. Dashed lines are a cubic
smoothing spline, and the two levels of the same networks are linked by
solid grey lines. \textbf{b}, Overall, interactions observed at the
local, regional, and continental scale have roughly equivalent
contributions to co-phylogenetic matching. Probability density was
smoothed using a Gaussian kernel density estimator. Raw probability
densities are shown as semi-transparent bars. Interaction-level
contributions are unitless, and lower values represent stronger
signal.\label{fig:contributions}}
\end{figure}

On the other hand, system-level differences say little about the
behavior of individual interactions. Despite the fact most
coevolutionary mechanisms act at the interaction level (Thompson 1999),
most \emph{measures} of it are expressed at the community level. We
observe here that networks with interactions that are important for
co-phylogenetic matching at the continental scale are also important for
co-phylogenetic matching at the local and regional scales as well
(\(\rho = 0.95\); \xrefname{fig.}\cref{fig:contributions}A).
Intriguingly, we also find that the distribution of individual
interactions' contributions to co-phylogenetic matching is strongly
conserved, regardless of the scale at which the interactions are
quantified (\xrefname{fig.}\cref{fig:contributions}B). Because
interactions differ between each other in terms of their total
contribution to co-phylogenetic matching, this implies that their
distribution across networks (\emph{i.e.} whether the local network
contains a sample of strongly contributing, or weakly contributing,
interactions) is what actually drives differences in overall
co-phylogenetic matching. As such, network-level co-phylogenetic
matching emerges directly from the properties of interactions and is not
a property of the network itself.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{figure/consistency-1.pdf}
\caption{Spatial consistency of an interaction and its contribution to
co-phylogenetic matching. Note that because \emph{PACo} gives low scores
to interactions with a strong contribution to co-phylogenetic matching,
the y-axis is reversed. Spatial consistency is defined as the
probability of observing an interaction between two species given that
they were observed to co-occur (the size of the point is proportional to
the number of times an interaction was observed). Although statistically
significant, there was no biologically meaningful relationship between
spatial consistency and an interaction's importance for co-phylogenetic
matching in the continental network (\(R^2 \approx 0.01\),
\(\rho = -0.1\), \(p \leq 10^{-5}\)).\label{fig:consistency}}
\end{figure}

Beyond their contribution to co-phylogenetic matching, interactions also
ultimately differ in how frequently they vary when the species involved
co-occur (Carstensen et al. 2014; Olito \& Fox 2015; Trøjelsgaard et al.
2015). This can happen, for example, when one of the partners is able to
forage for optimal resources (Betts et al. 2015). Once more, the
literature on host-parasite interactions assumes that the reason why
some interactions are more frequent is because they reflect a
significant past history of coevolution (Guimaraes et al. 2007;
Nieberding et al. 2010); that is, the ecological constraints emerge from
evolutionary conservatism. Using a weighted Pearson's correlation
between the interaction frequency, interaction contribution to
co-phylogenetic matching, and the number of observations of each
interaction as the weight, we observe that this is marginally true
(\(\rho \approx -0.11\). \(t \approx -5.09\) with weights;
\(\rho \approx -0.10\), \(t \approx -4.6\) without; both significant at
\(\alpha = 0.05\); \xrefname{fig.}\cref{fig:consistency}). Recall that
the \emph{negative} correlation here arises from the fact that high
interaction-level values in PACo means \emph{low} contribution to
co-phylogenetic signal. Nevertheless, the significance of this result
ought to be tempered by the fact that the \(R^2\) of both regressions is
close to \(0.01\). Consequently, the association between spatial
consistency and contribution to co-phylogenetic signal, while
statistically significant, explains so little variance of either
quantities that it is likely of negligible biological importance. This
implies that the spatial consistency of an interaction does not
necessarily reflect its evolutionary past, but rather (possibly) extant
ecological processes.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{figure/scales-1.pdf}
\caption{The contribution to co-phylogenetic matching of the interaction
between two species is maintained across scales. For every site (ranked
on the x-axis), we show the Pearson's correlation between
interaction-level values of co-phylogenetic matching in the continental
network and the same in the local network. The size of each point is
proportional to the size of the network, and correlations for all sites
are significant at \(\alpha = 0.05\) except for those falling in the
grey shaded area.\label{fig:scales}}
\end{figure}

Ultimately, co-phylogenetic matching varies across scale because of the
simultaneous variation of species' interactions \emph{and} communities'
phylogenetic tree structure. In a system characterized by substantial
turnover, we would expect the contribution of each separate interaction
to differ across scales as well. Instead, we observe here that
interactions that contribute strongly to co-phylogenetic matching at the
continental scale \emph{also} show a significant tendency to contribute
strongly at the local (\(p < 0.05\) for positive correlations in 48 out
of 51 networks) and regional (in 47 out of 51 networks), and this
observation is independent of network-wide co-phylogenetic matching
(\xrefname{fig.}\cref{fig:scales}). Remarkably, this result implies that
the remnants of co-phylogenetic inertia are still locally detectable in
\emph{individual interactions} even though shared evolutionary history
regularly fails to leave its imprint on most local networks.

@@

\hypertarget{ref-coel17nbp}{}
\textbf{Coelho et al.} (2017). Neutral biogeography of phylogenetically
structured interaction networks. \emph{Ecography.}:n/a--a.

@@

\hypertarget{ref-dev07wch}{}
\textbf{De Vienne et al.} (2007). When can host shifts produce congruent
host and parasite phylogenies? A simulation approach. \emph{Journal of
Evolutionary Biology.} 20:1428--38.
